---
mydisqus: "differences"
mytitle: "Comparing crypto prices between exchanges"
mysub: "Are there differences? How big? How often? on which exchange and currency?"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_crypto.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: TRUE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---


<div class="mycontent">





#Data
***
Prices of 5 cryto currencies have been harvested for 2 weeks on 5 different exchanges as described [here](). About 800,000 data points are available, allowing to compare prices between exchanges at a high resolution.

<br>

Here is an overview of the Etherum price on the period on the 5 exchanges:
```{r load_data, fig.align="center", warning=FALSE, message=FALSE, fig.width=10}
# Library
library(tidyverse)
library(DT)
library(plotly)
library(hrbrthemes)

# Load result
load("../DATA/public_ticker_harvest.Rdata")
Ticker$last <- as.numeric(Ticker$last)
Ticker= head(Ticker, 100000)

# Plot
p <- Ticker %>%
  filter(symbol=="ETHEUR") %>%
  ggplot( aes(x=time, y=last, color=platform, group=platform)) +
    geom_line() +
    theme_ipsum()
ggplotly(p)
```


*Note*: this graphic is interactive: zoom on a specific area and hover lines for more info.








#Differences {.tabset}
***
Can we try to quantify the difference? At each timestamp, we can calculate the difference between 2 exchanges, in euros and in %. We can then represent this difference. Here is an example with the difference between Bitstamp and Cex.
```{r show_differences, fig.width=10, fig.height=5, warning=FALSE}
plot_last_difference <- function(plat1, plat2, currency){

  # First I calculate the differnce of the 'last' price between both exchanges:
  diff <- Ticker %>%
    filter(symbol==currency) %>%
    filter(platform %in% c(plat1, plat2)) %>%
    select(time, platform, last, symbol) %>%
    spread(platform, last) %>%
    mutate(diff=.[[4]] - .[[3]], diff_perc=(.[[4]] - .[[3]]) / .[[4]] *100 ) %>%
    filter(!is.na(diff_perc))
  
  # Then I plot the result
  p <- ggplot(diff, aes(x=time, y=diff_perc, group=symbol, fill=symbol)) +
      geom_area(fill="#69b3a2") +
      facet_wrap(~symbol, nrow=5) +
      theme_ipsum() +
      theme(legend.position="none")
  
  ggplotly(p)
}

plot_last_difference("Kraken", "Bitstamp", "ETHEUR")
```

*Note*: it is important that this chart represent the price of the last transaction done. However, note that we buy crypto at the `ask` price, and sell it at the `bid` price. Differences are thus smaller than that in reality.




#- What we really get {.tabset}
***
It's important . For each time stamp, I can calculate the potential gain taking that into account:
```{r, fig.width=10, fig.height=8, warning=FALSE}
# A function that calculates the difference between 2 platforms for every currency at each time stamp
find_askbid_difference <- function(plat1, plat2){
  
  diff <- Ticker %>% 
    filter(platform %in% c(plat1, plat2)) %>%
    select(time, platform, symbol, ask, bid) %>%
    mutate(ask=as.numeric(ask), bid=as.numeric(bid)) %>%
    gather(temp, value, -time, -platform, -symbol) %>%
    mutate(platform=gsub(plat1,"plat1", platform)) %>%
    mutate(platform=gsub(plat2,"plat2", platform)) %>%
    unite(temp1, platform, temp, sep="_") %>%
    spread( key=temp1, value=value) %>%
    mutate( 
      diff1=(plat1_bid-plat2_ask)/plat1_bid*100, 
      diff2=(plat2_bid-plat1_ask)/plat2_bid*100
    ) %>%
    rowwise() %>%
    mutate( diff_perc=max(diff1, diff2) ) %>%
    filter(!is.na(diff_perc))
  return(diff)
}

plot_askbid_difference <- function(diff){
  p <- ggplot(diff, aes(x=time, y=diff_perc, group=symbol, fill=symbol)) +
        geom_area() +
        facet_wrap(~symbol, nrow=5) +
        theme(legend.position="none") 
    
  ggplotly(p)
} 
```






#Significant differences
***
What interest us is the number of time the difference between 2 platforms reaches a threshold that allow to perform arbitrage. Let's try to quantify how many times we reach this threshold for every currency between kraken and bitstamp.

```{r relationship_thres_numDiff, align="center", warning=FALSE}
# A function that counts the number of significant differences for several threshold
find_signif_diff <- function(diff){
  nbSignifDiff=data.frame()
  for( i in seq(0.7,4,0.2)){
    df <-  diff %>%
      group_by(symbol) %>%
      filter(diff_perc > i) %>%
      summarise( nb_over_thres = n() ) %>%
      mutate( thres = i) %>%
      arrange( nb_over_thres )
    nbSignifDiff <- rbind( nbSignifDiff, df)
  }
  return(nbSignifDiff)
}

# A function that plot this number of diff
plot_signif_diff <- function(nbSignifDiff){
  ggplot(nbSignifDiff, aes(x=thres, y=nb_over_thres, group=symbol, color=symbol)) +
    geom_line() +
    ylab("Number of cases in x days") +
    xlab("Difference threshold (%)")
}

# Apply that to a first example:
diff <- find_askbid_difference("Cex", "Bitstamp")
nbSignifDiff <- find_signif_diff(diff)
plot_signif_diff(nbSignifDiff)
```





Now, let's make this calculation for every pair of plateform.
```{r all_over_thres, align="center", warning=FALSE, fig.width=10}

# find all the pairs of platform
list_platforms <- Ticker$platform %>% unique() %>% as.character()
list_pairs <- combn(list_platforms, 2)

# find the number of significant differences for each pair
bilan=data.frame()
for(i in c(1:ncol(list_pairs))){
  diff <- find_askbid_difference(list_pairs[1,i], list_pairs[2,i])
  nbSignifDiff <- find_signif_diff(diff)
  nbSignifDiff$pair <- paste( list_pairs[1,i], list_pairs[2,i], sep="-")
  bilan <- rbind(bilan, nbSignifDiff)
}

#
ggplot(bilan, aes(x=thres, y=nb_over_thres, group=symbol, color=symbol)) +
  geom_line() +
  ylab("Number of cases in x days") +
  xlab("Difference threshold (%)") +
  facet_wrap(~pair, ncol=5)
```


#- Conclusion
***

- Based on these data, it looks like our best arbitrage opportunity takes place between Cex and Bitstamp.
- However, since Cex is more expensive than bitstamp on average, it is gonna to keep crypto on Cex and Euro on bitstamp
- BCH is the best option if we want to perform few transaction with high gain. XRP looks better for more transacation
</div>













