---
mydisqus: "private_API"
mytitle: "Trading crypto in command line using APIs"
mysub: "What is an API? How does it work? How to buy and sell crypto?"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_crypto.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: TRUE
      df_print: "paged"
      code_folding: "show"
      includes: 
        after_body: footer.html
---


<div class="mycontent">









#We need APIs
***
As explained in the previous posts, an arbitrage strategy has absolutely no chances of success if not `100% automatized`. Trading with your mouse through web interfaces would take way too long and arbitrage opportunities would be over even before you had the time to seize it.

<br>

Fortunately it is possible to exchange information with crypto exchanges through their API ([Application Programming Interface](https://en.wikipedia.org/wiki/Application_programming_interface)). Basically it allows to buy and sell crypto in command line, and thus to create a bot that do it automatically when a price difference is detected.




#Code resource
***

Let's suppose I want to know what is my balance on my Bitstamp account. It is possible to do so using the [R programming language](https://www.r-project.org) using the code below. Note that you first need to:

- Have a bitstamp account
- Set up the account to authorize access through API
- Get your codes

```{r, eval=FALSE}
# Some library are necessary to make these function work
library(dplyr)
library(digest)       # For the hmac function that allows to create the signature
library(RCurl)        # Pour récupérer les données en appelant des URLS
library(jsonlite)     # To go from JSON to data frame
  
# to access the Bitstamp API (enter your codes)
key_bitstamp="yyxxxxxxxxxxxxxxxxxxxxxxxxxxxxx"
secret_bitstamp="vfxxxxxxxxxxxxxxxxxxxxxxxxxxxx"
customer_id_bitstamp="zxxxxxx"
  
# A function that returns my Bitstamp balance
get_my_balance_bitstamp=function(){
    nonce=as.character(as.numeric(Sys.time()) * 1000000)
    signature = toupper( hmac(key = secret_bitstamp, object = paste0( nonce, customer_id_bitstamp, key_bitstamp), algo = "sha256"))
    post_data <- paste0("key=", key_bitstamp, "&signature=", signature, "&nonce=", nonce)
    curl <- getCurlHandle()
    query_result_json <- rawToChar(getURLContent(curl = curl, url = "https://www.bitstamp.net/api/v2/balance/", binary = TRUE, postfields = post_data))
    result=fromJSON(query_result_json,  flatten=TRUE) %>% data.frame()
    result=result[ , c("bch_available", "btc_available", "eth_available", "eur_available", "ltc_available", "xrp_available")]
    return(result)
  }
  
#Use the function
get_my_balance_bitstamp()
```

<br>

I've written a [set of function](https://github.com/holtzy/Cryp-To/tree/master/FUNCTIONS) allowing to `get balance`, `buy` crypto and `sell` crypto for Bitstamp, Cex, and Kraken. You can find these function [here](https://github.com/holtzy/Cryp-To/tree/master/FUNCTIONS). Here is an example of utilization:

```{r, eval=FALSE}
# Enter your bitstamp codes
key_bitstamp="yyxxxxxxxxxxxxxxxxxxxxxxxxxxxxx"
secret_bitstamp="vfxxxxxxxxxxxxxxxxxxxxxxxxxxxx"
customer_id_bitstamp="zxxxxxx"

# Source the script with bitstamp functions:
source("https://raw.githubusercontent.com/holtzy/Cryp-To/master/FUNCTIONS/Private_API_Functions_Bitstamp.R")

# Buy Etherum
from_euro_to_crypto_bitstamp(0.05, "ETHEUR" )
```




#Next step
***
Now that we have an arbitrage algorithm and a set of function allowing to buy and sell, let's try with real data!

<br>
<center>
<a class="btn btn-primary" href="https://holtzy.github.io/Crypto-Arbitrage/CONTENT/arbitrage.html">Let's play for real</a>
</center>
<br><br>


</div>













