---
mydisqus: "public_API"
mytitle: "Harvesting crypto currency prices"
mysub: "Recovering 1M data points on 5 exchanges and 5 currencies"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_crypto.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: TRUE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---


<div class="mycontent">









#Public exchange API
***
Crypto currencies are bought and sold on different `exchanges`. Basically it is like a bank, but for crypto. The price at which a token is traded depends on the offer and on the demand. Thus it evolves permanently, every couples of seconds.

<br>

It is possible to recover this price using the public API of exchanges. Let's say you want to know the `Bitcoin price` in the [Kraken]() exchange. You can do that in your browser, typing this URL:

[https://api.kraken.com/0/public/Ticker?pair=BTCEUR](https://api.kraken.com/0/public/Ticker?pair=BTCEUR)

It gives you several information, three being of interest:

- `last`: the price at the last transaction
- `bid`: price at which you can sell your crypto
- `ask`: price at which you can buy a crypto





#Do it programming
***

It is totally possible to do the same `programming`. This is handy since it will allow to recover the prices every couple of second automatically.  

Here is an example using the R programming language to get the price of the `bitcoin` on [Kraken]() and showing it in a clean table:

```{r, warning=FALSE, message=FALSE, eval=TRUE}
# package
library(tidyverse)

# Recover the information
library(RCurl)
adress <- "https://api.kraken.com/0/public/Ticker?pair=BTCEUR"
ticker <- getURLContent(adress)

# Make the format more readable
require(jsonlite)
tmp <- fromJSON(ticker)$result[[1]]
result = data.frame(ask=tmp$a[1], bid=tmp$b[1], last=tmp$c[1], open=tmp$o, low=tmp$l[1], high=tmp$h[1], volume=tmp$v[1], volumeQuote=NA, timestamp=NA)

# Show result
library(kableExtra)
options(knitr.table.format = "html")

tmp <- result %>% 
  select(-volumeQuote, -timestamp)
apply(tmp, 2, function(x){round(as.numeric(x),0)}) %>%
  t() %>%
  kable() %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)
```



#Code resource
***

I've written a [set of functions](https://github.com/holtzy/Cryp-To/blob/master/FUNCTIONS/Public_Market_Functions.R) allowing to get the price of many different currencies for the 5 main exchanges. You can easily use these functions. For instance, type the code below in R:

<div style="background-color:#F8F8F8; border-radius: 1px; padding: 20px;">


 # Source de functions that are stored on github  
 source("https://raw.githubusercontent.com/holtzy/Cryp-To/master/FUNCTIONS/Public_Market_Functions.R")

 # Use it: price of the bitcoin on bitstamp  
 get_bitstamp(time, "BTCEUR")


</div>




#Harvesting (a lot of) data
***

I've harvested crypto prices:

- *for 2 weeks* - between the `5th and the 19th of February`.
- *for 5 currencies* - bitcoin (BTC), bitcoin cash (BCH), etherum (ETH), litecoin (LTC) and ripple (XRP)
- *for 5 exchanges* - Coinbase, Kraken, Bitstamp, Bitfinex and Cex.io


This was easily done using an infinite loop that called the functions described above. The exact script used for this work is available [here](). 800,000 data points were recovered.

<br>


The resulting dataset is available on [github]() in a compressed format. You can easily read it in R doing:

<div style="background-color:#F8F8F8; border-radius: 1px; padding: 20px;">

 load("../DATA/public_ticker_harvest.Rdata")  
 head(Ticker)

</div>

<br>

As a teaser, here is the evolution of the etherum price on Bitstamp on this period of time:

```{r, fig.align="center", warning=FALSE, cache=TRUE, fig.width=9, message=FALSE, fig.height=4}
load("../DATA/public_ticker_harvest.Rdata")

library(hrbrthemes)
Ticker %>%
  filter( symbol == "ETHEUR" ) %>%
  filter(platform == "Bitstamp") %>%
  ggplot( aes(x=time, y=as.numeric(last))) +
    geom_ribbon(aes(ymin=450, ymax=as.numeric(last)),  fill="#69b3a2", color="transparent", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ggtitle("Evolution of Etherum price on the period") +
    ylab("bitcoin price ($)") +
    theme_ipsum() +
    theme(
      plot.title = element_text(size=12)
    )
```

Next step: [analysing this dataset]().


</div>













