---
title: "Public API"
author: "by [Yan Holtz](https://github.com/holtzy/) - `r format(Sys.time(), '%d %B %Y')`"
---

<br><br>



> What is the cheapest Exchange to buy Crypto?


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(rmarkdown)    # You need this library to run this template.
library(epuRate)      # Install with devtools: install_github("holtzy/epuRate", force=TRUE)
library(DT)
```


```{r load_data}
load("DATA/public_ticker_harvest.Rdata")
Ticker$last <- as.numeric(Ticker$last)
```












#- What is the most expensive exchange?
Is it always the same exchange that is more expensive? Or does it swap? Let's find out!

For bitcoin
```{r swap, fig.width=10}
tmp <- Ticker %>%
  filter(symbol=="BTCEUR") %>%
  select(time, platform, last) %>% 
  spread( platform, last) %>%
  rowwise() %>%
  mutate( mymean = mean(c(coinbase, Kraken, Bitstamp, Bitfinex, Cex), na.rm=T)) %>%
  gather( platform, value, -c(1,7)) %>% 
  mutate( diff=value-mymean) 

tmp %>%
  mutate(platform=as.factor(platform)) %>%
  na.omit() %>%
  ggplot( aes(x=fct_reorder(platform, diff), y=diff, color=platform)) +
    geom_boxplot() #outlier.shape = NA) #+
    #coord_cartesian(ylim = quantile(tmp$diff, c(0.1, 0.9), na.rm=T))
```

This graph shows us several information.
One of the information is that we can have somteimes really huge difference.
Let's zoom on a huge difference, where CEX is about 200$ more expensive,

```{r , fig.width=10}

data = Ticker %>% 
    filter(symbol=="BTCEUR") %>%
    select(time, platform, last) %>%
    filter(time>1517890104 ) %>%
    spread(platform, last)
  
  # Go to a xts format
  library(xts)
  don=xts( x=data[,-1], order.by=data$time)

  #time_series <- xts(tmp)
  library(dygraphs)
  dygraph(don) %>%
    dyLegend(show = "always", hideOnMouseOut = FALSE) %>%
    #dyRoller(rollPeriod = 5) %>%
    dyRangeSelector()
```





For all other crypto
```{r  fig.width=10}
tmp <- Ticker %>%
  select(time, platform, last, symbol) %>% 
  spread( platform, last) %>%
  rowwise() %>%
  mutate( mymean = mean(c(coinbase, Kraken, Bitstamp, Bitfinex, Cex), na.rm=T)) %>%
  gather( platform, value, -c(1,2,8)) %>% 
  mutate( diff= (value-mymean)/mymean*100) 

tmp %>%
  filter(diff>-2 & diff<2) %>%
  mutate(platform=as.factor(platform)) %>%
  na.omit() %>%
  ggplot( aes(x=fct_reorder(platform, diff), y=diff, fill=platform)) +
    geom_violin(alpha=0.8) +
    facet_wrap(~symbol, scale="free_y") +
    xlab("") +
    theme(legend.position="none")
```













