---
title: "Comparing Crypto currency platforms"
author: "by [Yan Holtz](https://github.com/holtzy/) - `r format(Sys.time(), '%d %B %Y')`"
output:
  epuRate::epurate:
    toc: FALSE
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



<br><br>


> Crypto currencies are quite trendy those days. Definition. You can buy some in every places. But is the price the same everywhere?

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(rmarkdown)    # You need this library to run this template.
library(epuRate)      # Install with devtools: install_github("holtzy/epuRate", force=TRUE)
library(DT)
```




#- Call a platform API
***
Let's say you want to harvest the Bitcoin in the Kraken plateform. You have to call it's public API. You can do that in your browser, typing this URL:  
[https://api.kraken.com/0/public/Ticker?pair=BTCEUR]()

Description of the result.

Now let's do the same using the R programming language:

```{r, warning=FALSE, message=FALSE}
# Recover the information
library(RCurl)
adress <- "https://api.kraken.com/0/public/Ticker?pair=BTCEUR"
ticker <- getURLContent(adress)

# Make the format more readable
require(jsonlite)
tmp <- fromJSON(ticker)$result[[1]]
result = data.frame(ask=tmp$a[1], bid=tmp$b[1], last=tmp$c[1], open=tmp$o, low=tmp$l[1], high=tmp$h[1], volume=tmp$v[1], volumeQuote=NA, timestamp=NA)
#datatable(result, rownames = FALSE, options = list(dom=TRUE) )
result
```

We can transform this piece of code in a function.
```{r}
get_kraken=function(time, currency){
  adress=paste("https://api.kraken.com/0/public/Ticker?pair=", currency, sep="" )
  tmp=getURLContent(adress) %>% fromJSON() 
  tmp=tmp$result[[1]]
  result = data.frame(time=time, platform="Kraken", ask=tmp$a[1], bid=tmp$b[1], last=tmp$c[1], open=tmp$o, low=tmp$l[1], high=tmp$h[1], volume=tmp$v[1], volumeQuote=NA, timestamp=NA, symbol=currency)
  return(result)
}
```







#- Harvest 5 plateforms for 5 currencies {.tabset .tabset-fade .tabset-pills}
***
I've created a text file with one function like the one above per plateform. The file is located in my [github](https://github.com/holtzy/Cryp-To/blob/master/FUNCTIONS/Public_Market_Functions.R). We can source it doing:

```{r, warning=FALSE, message=FALSE}
# POTENTIAL PROBLEM WITH THE LOADING OF LIBRARIES
source("https://raw.githubusercontent.com/holtzy/Cryp-To/master/FUNCTIONS/Public_Market_Functions.R")
```


Now we are ready for an infinite loop! I'm gonna run these functions for several platforms and several crypto every 10 secondes approx.
```{r, eval=FALSE}
#Initialize an emtpy result table
Ticker=as.data.frame(matrix(NA, 0, 12))
names(Ticker) = c("time", "platform", "ask", "bid", "last", "open", "low", "high", "volume", "volumeQuote", "timestamp", "symbol")

# Run the analysis
num=0
while(TRUE){
  
  # Keep a trace of where we are in this loop
  num=num+1
  cat("Let's start the iteration number: ",num, " at ", Sys.time(), "\n" )
  
  # BITCOIN (BTC)
  time=Sys.time()
  tmp=try(get_coinbase(time, "BTCEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_kraken(time, "BTCEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_bitstamp(time, "BTCEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_bitfinex(time, "BTCEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_cex(time, "BTCEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  
  # ETHERUM (ETH)
  time=Sys.time()
  tmp=try(get_coinbase(time, "ETHEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_kraken(time, "ETHEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_bitstamp(time, "ETHEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_cex(time, "ETHEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  
  # LITECOIN (LTC)
  time=Sys.time()
  tmp=try(get_coinbase(time, "LTCEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_kraken(time, "LTCEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_bitstamp(time, "LTCEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }

  # BITCOINCASH (BCH)
  time=Sys.time()
  tmp=try(get_coinbase(time, "BCHEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_kraken(time, "BCHEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_bitstamp(time, "BCHEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_cex(time, "BCHEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  
  # RIPPLE (XRP)
  time=Sys.time()
  tmp=try(get_kraken(time, "XRPEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_bitstamp(time, "XRPEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }
  tmp=try(get_cex(time, "XRPEUR")) ; if(is.data.frame(tmp)){ Ticker=rbind(Ticker, tmp) }

  # Every ten loop I save the file
  if (num %% 1 == 0){  save(Ticker, file="../DATA/public_ticker_harvest.Rdata")  }

}
```

I've run this script during X days from the to the. This dataset is called `public_ticker_harvest.Rdata` that is available [on github](https://github.com/holtzy/Cryp-To/tree/master/DATA).



#- Visualize currency values {.tabset .tabset-fade .tabset-pills}
***
Let's visualize the differences!

```{r load_data}
# Load result
load("../DATA/public_ticker_harvest.Rdata")
Ticker$last <- as.numeric(Ticker$last)
#Ticker <- Ticker %>% sample_n(200000)

# Occurence of each
table(Ticker$platform, Ticker$symbol)
```

A function that make a plot for a specific currency:
```{r function_plot_a_currency}
plot_ticker=function(currency){
  
  # I need a wide format for Dygraph
  data = Ticker %>% 
    filter(symbol==currency) %>%
    select(time, platform, last) %>%
    spread(platform, last)
  
  # Go to a xts format
  library(xts)
  don=xts( x=data[,-1], order.by=data$time)

  #time_series <- xts(tmp)
  library(dygraphs)
  dygraph(don) %>%
    dyLegend(show = "always", hideOnMouseOut = FALSE) %>%
    #dyRoller(rollPeriod = 5) %>%
    dyRangeSelector()
}
```




##BTC
```{r show_btc, message=FALSE, warning=FALSE, fig.width=10, fig.align="center"}
plot_ticker("BTCEUR")
```

##ETH

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.align="center"}
plot_ticker("ETHEUR")
```

##LTC
```{r, message=FALSE, warning=FALSE, fig.width=10, fig.align="center"}
plot_ticker("LTCEUR")
```

##XRP
```{r, message=FALSE, warning=FALSE, fig.width=10, fig.align="center"}
plot_ticker("XRPEUR")
```

##BCH
```{r, message=FALSE, warning=FALSE, fig.width=10, fig.align="center"}
plot_ticker("BCHEUR")
```





#- Difference

Can we try to quantify the difference?
```{r show_differences, fig.width=16, fig.height=8, warning=FALSE}
diff <- Ticker %>%
  filter(platform %in% c("Bitstamp", "Cex")) %>%
  group_by(time, symbol) %>%
  summarise(max=max( as.numeric(last) ), min=min( as.numeric(last) ) ) %>%
  mutate( diff=max-min, diff_perc=(max-min)/max*100 )

ggplot(diff, aes(x=time, y=diff_perc, group=symbol, fill=symbol)) +
    geom_area() +
    facet_wrap(~symbol, nrow=5) +
    theme(legend.position="none") +
    ylim(0,4)
```




```{r, fig.width=16, fig.height=8, warning=FALSE}
# A function that calculates the difference between 2 platforms for every currency at each time stamp
find_differences <- function(plat1, plat2){
  diff <- Ticker %>% 
    filter(platform %in% c(plat1, plat2)) %>%
    select(time, platform, symbol, ask, bid) %>%
    mutate(ask=as.numeric(ask), bid=as.numeric(bid)) %>%
    gather(temp, value, -time, -platform, -symbol) %>%
    mutate(platform=gsub(plat1,"plat1", platform)) %>%
    mutate(platform=gsub(plat2,"plat2", platform)) %>%
    unite(temp1, platform, temp, sep="_") %>%
    spread( key=temp1, value=value) %>%
    mutate( 
      diff1=(plat1_bid-plat2_ask)/plat1_bid*100, 
      diff2=(plat2_bid-plat1_ask)/plat2_bid*100
    ) %>%
    rowwise() %>%
    mutate( diff_perc=max(diff1, diff2) )
  return(diff)
} 

# Find differences between kraken and bitstamp
diff <- find_differences("Bitstamp", "Cex")

ggplot(diff, aes(x=time, y=diff_perc, group=symbol, fill=symbol)) +
    geom_area() +
    facet_wrap(~symbol, nrow=5) +
    theme(legend.position="none") +
    ylim(0,4)
```


# - Cryptocurrencies are correlated?
Can I compare the evolution of several crypto on the same plot?
```{r}
Ticker %>% 
  filter(platform=="Kraken") %>%
  ggplot( aes(x=time, y=last, color=symbol)) +
    geom_line() +
    facet_wrap(~symbol, scale="free", ncol=1) +
    theme( legend.position="none")
```

Another way to look at that is to compare % of max value for each crypto
```{r, fig.width=12}
Ticker %>% 
  filter(platform=="Cex") %>%
  group_by(symbol) %>%
  mutate(last_normalized = last/mean(last)) %>%
  ggplot( aes(x=time, y=last_normalized, color=symbol)) +
    geom_line()
```
This is really important. It means that we can do another type of arbitrage. Using 2 crypto and considering that the price of 1 crypto will necessarily follow the price of a second one, sooner or later. When one is late, we need to buy some, it will increase one day.


# - Significant difference
What interest us is the number of time the difference between 2 platforms reaches a threshold that allow to perform arbitrage. Let's try to quantify how many times we reach this threshold for every currency between kraken and bitstamp.
```{r relationship_thres_numDiff, align="center", warning=FALSE}

# A function that counts the number of significant differences for several threshold
find_signif_diff <- function(diff){
  nbSignifDiff=data.frame()
  for( i in seq(0.7,3,0.1)){
    df <-  diff %>%
      group_by(symbol) %>%
      filter(diff_perc > i) %>%
      summarise( nb_over_thres = n() ) %>%
      mutate( thres = i) %>%
      arrange( nb_over_thres )
    nbSignifDiff <- rbind( nbSignifDiff, df)
  }
  return(nbSignifDiff)
}

nbSignifDiff <- find_signif_diff(diff)

ggplot(nbSignifDiff, aes(x=thres, y=nb_over_thres, group=symbol, color=symbol)) +
  geom_line() +
  ylab("Number of cases in x days") +
  xlab("Difference threshold (%)")
```

Now, let's make this calculation for every pair of plateform.
```{r all_over_thres, align="center", warning=FALSE, fig.width=10}

# find all the pairs of platform
list_platforms <- Ticker$platform %>% unique() %>% as.character()
list_pairs <- combn(list_platforms, 2)

#
bilan=data.frame()
for(i in c(1:ncol(list_pairs))){
  diff <- find_differences(list_pairs[1,i], list_pairs[2,i])
  nbSignifDiff <- find_signif_diff(diff)
  nbSignifDiff$pair <- paste( list_pairs[1,i], list_pairs[2,i], sep="-")
  bilan <- rbind(bilan, nbSignifDiff)
}

#
ggplot(bilan, aes(x=thres, y=nb_over_thres, group=symbol, color=symbol)) +
  geom_line() +
  ylab("Number of cases in x days") +
  xlab("Difference threshold (%)") +
  facet_wrap(~pair, ncol=5)
```










#- What is the most expensive exchange?
Is it always the same exchange that is more expensive? Or does it swap? Let's find out!

For bitcoin
```{r swap, fig.width=10}
tmp <- Ticker %>%
  filter(symbol=="BTCEUR") %>%
  mutate(Bitfinex=replace(last, last<1000, NA)) %>%
  select(time, platform, last) %>% 
  spread( platform, last) %>%
  rowwise() %>%
  mutate( mymean = mean(c(coinbase, Kraken, Bitstamp, Bitfinex, Cex), na.rm=T)) %>%
  gather( platform, value, -c(1,7)) %>% 
  mutate( diff=value-mymean) 

tmp %>%
  mutate(platform=as.factor(platform)) %>%
  na.omit() %>%
  ggplot( aes(x=fct_reorder(platform, diff), y=diff, color=platform)) +
    geom_boxplot()
```

This graph shows us several information.
One of the information is that we can have somteimes really huge difference.
Let's zoom on a huge difference, where CEX is about 200$ more expensive,

```{r swap, fig.width=10}

data = Ticker %>% 
    filter(symbol=="BTCEUR") %>%
    select(time, platform, last) %>%
    filter(time>1517890104 ) %>%
    spread(platform, last)
  
  # Go to a xts format
  library(xts)
  don=xts( x=data[,-1], order.by=data$time)

  #time_series <- xts(tmp)
  library(dygraphs)
  dygraph(don) %>%
    dyLegend(show = "always", hideOnMouseOut = FALSE) %>%
    #dyRoller(rollPeriod = 5) %>%
    dyRangeSelector()
```















